# Exercise 3.2

## Find the the log-likelihood function

The log-likelihood function of this distribution is

$$ \ell(\mathbf{x}, \theta) = \sum_{i=1}^n \log\{1-\cos(x_i-\theta)\} - n\log2\pi$$


```{r}
x <- c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
       2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)

loglikelihood <- function(theta){
  n <- length(x)
  s <- sum(log(1 - cos(x - theta))) + n * log(2 * pi)
  return(s)
}
loglikelihood <- Vectorize(loglikelihood)
curve(loglikelihood, -pi, pi, xlab = expression(theta), ylab = "log-likelihood")
```

## Find the method-of-moments estimator

The expectation of $\mathbf{x}|\theta$ is
\begin{align*}
\mathbb E (x | \theta) &= \int_{0}^{2\pi} x \frac{1-\cos(x-\theta)}{2\pi} dx \\
&= \frac{1}{2\pi} \int_{0}^{2\pi} x - x\cos(x-\theta) dx \\
&= \pi + \sin(\theta) \\
&= \bar{X_n}
\end{align*}

Thus, 

```{r}
theta_tilde <- asin(mean(x) - pi)
theta_tilde
```

## Find the MLE 

Since

$$\frac{\partial\ell(\mathbf{x}; \theta)}{\partial\theta} = \sum_{i=1}^n \frac{-\sin(x_i - \theta)}{1-\cos(x_i-\theta)}$$

$$\frac{\partial^2\ell(\mathbf{x}; \theta)}{\partial \theta^2} = \sum_{i=1}^n \frac{\cos(x_i-\theta) - \cos^2(x_i-\theta)-\sin^2(x_i-\theta)}{(1-\cos(x_i-\theta))^2}$$

The Newton-Raphson method is

$$\hat\theta^{(t+1)} = \hat\theta^{(t)} - \left\{\frac{\partial^2\ell(\mathbf{x}; \hat\theta^{(t)})}{\partial \theta^2}\right\}^{-1}\frac{\partial\ell(\mathbf{x}; \hat\theta^{(t)})}{\partial\theta}$$

```{r}
lfd <- function(theta){
  sum(-sin(x-theta)/(1-cos(x-theta)))
}

lsd <- function(theta){
  sum((cos(x-theta) - (cos(x-theta))^2 - (sin(x-theta))^2)/(1-cos(x-theta))^2)
}

Newton <- function(init){
  theta0 <- init
  i <- 0
  diff <- 1
  msg <- "converge"
  while(abs(diff) > 0.0000001){
    lfd <- lfd(theta0)
    lsd <- lsd(theta0)
    diff <- (lfd/lsd)
    theta1 <- theta0 - diff
    theta0 <- theta1
    i <- i+1
    #cat(i)
    if(i >= 150){
      msg <- "Not converge"
      theta0 <- Inf
      break
    }
  }
  return(list(theta = theta0, itr = i, msg = msg))
}
Newton(theta_tilde)
```

## $\theta_0 = 2.7$ or $\theta_0 = -2.7$

```{r}
Newton(-2.7)
Newton(2.7)
```

The $\hat\theta$ we got is different.

## Repeat the above using 200 equally spaced starting values

```{r}
init <- seq(-pi, pi, length.out=200)
result <- NULL
for(initi in init){
  result <- rbind(result, c(initi, Newton(initi)$theta))
}
colnames(result) <- c("Initial_value", "theta_hat")
split(result, result[,2])
```
