# Exercise 3.3

## Fit the population growth model to the beetles data using the Gauss-Newton approach

```{r HW4_4, warning=FALSE}
library(graphics)
library(Matrix)
beetles <- data.frame(
    days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
    beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))

##' define the sum of squared errors function
sqerr <- function(k, r) {
  s <- matrix(0, nrow = length(k), ncol = length(r))
  for (i in 1:length(k)) {
    for (j in 1:length(r)) {
      s[i, j] <- sum((beetles$beetles - k[i] * beetles$beetles[1] / 
              (beetles$beetles[1] + (k[i] - beetles$beetles[1]) * 
                 exp(-r[j]*beetles$days)))^2)
    }
  }
  s
}

##' define z function
z.vec <- function(k, r) {
  n <- length(beetles$days)
  z <- rep(0, n)
  for (i in 1:n) {
   z[i] <- beetles$beetles[i] - k*beetles$beetles[1] / 
     (beetles$beetles[1] + (k - beetles$beetles[1])*exp(-r*beetles$days[i])) 
  }
  return(z)
}

##' define A matrix
A.mat <- function(k, r) {
  n <- length(beetles$days)
  A <- matrix(0, nrow = n, ncol = 2)
  for (i in 1:n) {
    A[i, 1] <- beetles$beetles[1]^2 * (1-exp(-r*beetles$days[i])) / 
      (beetles$beetles[1] + (k-beetles$beetles[1])*exp(-r*beetles$days[i]))^2
    
    A[i, 2] <- beetles$beetles[1]*k*beetles$days[i]*(k-beetles$beetles[1]) * 
      exp(-r*beetles$days[i]) / (beetles$beetles[1]+(k-beetles$beetles[1]) * 
                                   exp(-r*beetles$days[i]))^2
  }
  return(A)
}

gaussNewton.beetles <- function(para0, z.vec, A.mat, maxiter = 100, 
                                tol = .Machine$double.eps^0.5) {
  para <- para0
  for (i in 1:maxiter) {
    Amat <- A.mat(para[1], para[2])
    zvec <- z.vec(para[1], para[2])
    para.new <- para + solve(t(Amat) %*% Amat + 0.0001*diag(nrow = 2)) %*% 
      t(Amat) %*% zvec
    if (sum(abs(para.new - para)) < tol) break
    para <- para.new
  }
  if (i == maxiter) warning("maximum iteration has reached")
  return(list(root = para, niter = i))
}

fit <- gaussNewton.beetles(para0 = c(1200, 0.1), z.vec = z.vec, A.mat = A.mat)
fit$root

k <- seq(100, 1500, by = 10)
r <- seq(0.1, 0.5, by = 0.001)
z <- sqerr(k, r)
contour(k, r, z, xlab = "k", ylab = "r", main = "contour plot for squared error")
```

Estimation using Gauss-Newton method is $\hat{k} = 1049$, $\hat{r} = 0.12$. 

## Log normal assumption

```{r HW4_5, warning=FALSE}
##' define log-likelihood function
loglikeli <- function(para) {
  k <- para[1]
  r <- para[2]
  sigma2 <- para[3]
  l <- sum(-log(2*pi*sigma2)/2 - (log(beetles$beetles) - log(k) - 
                                    log(beetles$beetles[1]) + 
                                    log(beetles$beetles[1] + 
                                          (k-beetles$beetles[1]) * 
                                          exp(-r*beetles$days))^2)/2/sigma2)
  return(l)
}

##' define gradiant of loglikelihood function
grad.my <- function(para) {
  k <- para[1]
  r <- para[2]
  sigma2 <- para[3]
  g <- rep(0, 3)
  g[1] <- sum(-2*(log(beetles$beetles)-log(k)-log(beetles$beetles[1])+
                    log(beetles$beetles[1]+(k-beetles$beetles[1])*
                          exp(-r*beetles$days)))*
                (-1/k+exp(-r*beetles$days)/
                   (beetles$beetles[1]+
                      (k-beetles$beetles[1])*exp(-r*beetles$days)))/2/sigma2)
  g[2] <- sum(2*(log(beetles$beetles)-log(k)-log(beetles$beetles[1])+
                   log(beetles$beetles[1]+(k-beetles$beetles[1])*
                         exp(-r*beetles$days)))*
                beetles$days*(k-beetles$beetles[1])*exp(-r*beetles$days)/
                (beetles$beetles[1]+(k-beetles$beetles[1])*exp(-r*beetles$days))/
                2/sigma2)
  g[3] <- sum(-1/2/sigma2+(log(beetles$beetles)-log(k)-log(beetles$beetles[1])+
                             log(beetles$beetles[1]+(k-beetles$beetles[1])*
                                   exp(-r*beetles$days)))^2/2/sigma2^2)
  return(g)
}

fit <- constrOptim(theta = c(10, 0.1, 1), f = loglikeli, grad = grad.my, 
                   ui = diag(1, 3), ci = rep(0, 3), 
                   control = list(fnscale = -1), hessian = TRUE)
fit$par
fit$convergence
Diag(-solve(fit$hessian))
```

Using BFGS methods, set constrain to make $k$, $r$, $sigma^2$ nonnegative, we 
get the MLE estimates $\hat{k} = 103.98$, $\hat{r} = 12.12$, $\hat{\sigma}^2 = 2.91$. 
Using inverse of negative hessian, we have these estimates' variance to be 
$2.7\times 10^3$, $1.2\times 10^5$, $2.8$ respectively.

But there's problem that results will change dramatically with initial values. 
I just tried several different initial values, get the estimates and compare 
their corresponding loglikelihood. I just choose the one with maximum loglikelihood.